---
title: "Lab 8 - Linear Regression in R"
author: "Priscilla Hare"
date: "November 20, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Step 1: load the tidyverse package and hte 'salmon_imports.csv' file
```{r}

library(tidyverse)
salmon <- read_csv('salmon_imports.csv')

```


##Make an exploratory scatter plot
```{r}

salmon_scatter <- salmon %>% 
  ggplot(aes(x = year, y = salmon_fresh))+
  geom_point()

salmon_scatter

#Overall it looks linear but with some variation. Concernsa re the few spikes
```

##Linear regression:
-perform linear regression
-call the model name
-write out the model equation including actual variables and coefficients
-think about the model
```{r}
# format for linear regression is 
# model_name <- lm(y_variable ~ x_variable, data - df_name)

salmon_model <- lm(salmon_fresh ~ year, data = salmon)
salmon_model

# imports(tons) = -14982940 + 7550 (year)

```
####What does the slope mean?
The slope tells us how much slamon imports increase by each year
####What does the y -intercept mean. Why is it concerning? What does this mean about extrapolating this model for past values?
The y - intercept means that at year = 0 there is a negative salmon import. This does not make sense. For extrapolating for past values a limit on how fat back needs to be set so the model makes sense. 

##Model Diagnostics
-use plot(model_name) to view model diagnostics
-do residuals appear normally distirbuted
-any concerns about heteroscedasticity or outliers
```{r}
par(mfrow = c(2,2)) #use to view all graphs from plot at once
plot(salmon_model)

#residauls appear noramlly distrubuted
# some concerns about heteroscedasticitity
#maybe a few outliers
```

##Explore fit and model signigicance
-does year significantly predict salmon imports
-what does the R^2^ value mean in words
-what proportion of the variance in salmon imports is NOT explained by year
```{r}
summary(salmon_model)

#p < .001 
```
Year significantly predicts salmon imports
The R^2^ value means that 94.6% of the variance in salmon imports can be predicted by the year.
5.4% is not explained by year

##Prepare a final graph
```{r}

salmon_final_graph <- salmon %>% 
  ggplot(aes(x = year, y = salmon_fresh))+
  geom_point()+
  geom_smooth(method = lm, se = TRUE, size = 0.5, color = "gray20") + #plots the linear model with a confidence interval (se)
  theme_bw() +
  labs(x = "Year", y = "U.S. Annual Fresh Salmon Imports\n(1,000 pounds)", title = "Salmon Imports Model") +
  scale_x_continuous(expand = c(0,0), limits = c(1990, 2017))

salmon_final_graph
```

##Make predictions using the model
Lets say we want to make predictions for the years 2022, 2024, 2026, and 2028
```{r}

new_years <- data.frame(year = c(2022, 2024, 2026, 2028)) #creates a new data frame with thte years we want to predict using our model

#now feed this new data frame into the predict() function 

future_predict <- predict(salmon_model, newdata = new_years, interval = "confidence")

future_predict

predictions <- data.frame(new_years, future_predict)

predictions  #a new data frame that binds the predictions to the years

```

##Find Pearson's r (correlation)
```{r}

salmon_cor <- cor.test(salmon$year, salmon$salmon_fresh)

salmon_cor

#r=.974...strong positve correlation
```

##Concluding Statement

Year significantly predicts salmon imports (b = 7.550e+03, t(26) = 21.72, p < 0.001) with a strong postive correlation between the two (Pearson's r = 0.974). The overall model (imports(tons) = -14982940 + 7550 (year)) explains a significant amoint of variacne in imports (F(1,26) = 472, p < 0.001, R^2^ = 0.95).